---
title: "IMDB Movie Analysis"
output: html_document
params:
   min_year: 2005
   max_year: 2016
   imdb_score : 7
   facebook_thresh : 1000
---

```{r setup, include=FALSE}
library(dplyr)

knitr::opts_chunk$set(echo = FALSE)
df <-
read.csv(
"/Users/valantrahome/projects/data_science-r/movie_metadata2.csv",
header = TRUE,
stringsAsFactors = FALSE
)

```


This my first analysis of IMDB data using dplyr and published via _R Markdown_

####Number of Movies and average ratings by year####
```{r data1}
movies_rating_byCountryyear <-
  df %>% group_by(country, title_year) %>% summarize(
  imdb_score = round(mean(imdb_score), 2),
  facebook_likes = round(mean(movie_facebook_likes)),
  movie_cnt = n()
  ) %>% arrange(desc(title_year), desc(imdb_score))
  
  movies_rating_byCountryyear
```

####Number of movies by country and by year having a IMDB rating of `r paste(params$imdb_score)` or greater and average facebook likes for movies, directors and main actor####
_Only movies made between year `r paste(params$min_year)` and year `r paste(params$max_year)` are considered_

```{r data}
movies_country_title = df %>% group_by(country, title_year, imdb_score)  %>%
  summarise(
  movie_likes = mean(movie_facebook_likes),
  director_likes = mean(director_facebook_likes),
  actor1_likes = mean(actor_1_facebook_likes),
  movies = n()
  )  %>%
  filter(
  title_year >= params$min_year ,
  title_year <= params$max_year ,
  imdb_score >= params$imdb_score
  )  %>%
  filter(
  movie_likes > params$facebook_thresh,
  director_likes > params$facebook_thresh ,
  actor1_likes > params$facebook_thresh
  ) %>%
  arrange(desc(title_year))
  
  #publish data
  movies_country_title
```

###Countries by year, count of movies having rating more than _`r paste(params$imdb_score)`_ ###
```{r data2}
library(ggplot2)
ggplot(movies_country_title,
aes(x =  country , y = title_year, size = movies)) + geom_point()
```