---
title: "IMDB Movie Analysis"
output: html_document
params:
  min_year: 2005
  max_year: 2016
  imdb_score : 7
  movie_count : 5
  facebook_thresh : 1000
---

```{r setup, include=FALSE}
library(dplyr)

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
df <-read.csv(
  "./movie_metadata2.csv",
  header = TRUE,
  stringsAsFactors = FALSE
)

```


This my first analysis of IMDB data using dplyr and published via _R Markdown_

####Number of Movies and average ratings by year####
```{r data1, message=FALSE, warning=FALSE}
movies_rating_byCountryyear <-
  df %>% group_by(country, title_year) %>% summarize(
    imdb_score = round(mean(imdb_score), 2),
    facebook_likes = round(mean(movie_facebook_likes)),
    movie_cnt = n()
  ) %>% arrange(desc(title_year), desc(imdb_score))

movies_rating_byCountryyear
```

####Number of movies by country and by year having a IMDB rating of `r paste(params$imdb_score)` or greater and average facebook likes for movies, directors and main actor####
_Only movies made between year `r paste(params$min_year)` and year `r paste(params$max_year)` are considered_

```{r data, message=FALSE, warning=FALSE}
movies_country_title = df %>% group_by(country, title_year, imdb_score)  %>%
  summarise(
    movie_likes = mean(movie_facebook_likes),
    director_likes = mean(director_facebook_likes),
    actor1_likes = mean(actor_1_facebook_likes),
    movies = n()
  )  %>%
  filter(
    title_year >= params$min_year ,
    title_year <= params$max_year ,
    imdb_score >= params$imdb_score
  )  %>%
  filter(
    movie_likes > params$facebook_thresh,
    director_likes > params$facebook_thresh ,
    actor1_likes > params$facebook_thresh
  ) %>%
  arrange(desc(title_year))

#publish data
movies_country_title
```


```{r plotdata, message=FALSE, warning=FALSE}
library("ggplot2")
library("dplyr")
library("RColorBrewer")

plot_data <-
df %>% group_by(country, title_year) %>% summarize(
imdb_score = round(mean(imdb_score), 2),
a_facebook_likes = round(quantile(na.rm = TRUE, actor_1_facebook_likes, c(.90))),
m_facebook_likes = round(quantile(na.rm = TRUE, movie_facebook_likes, c(.90))),
d_facebook_likes = round(quantile(na.rm = TRUE, director_facebook_likes, c(.90))),

movie_cnt = n()
) %>% arrange(desc(title_year), desc(imdb_score)) %>%
filter(
  title_year >= params$min_year ,
  title_year <= params$max_year ,
  imdb_score >= params$imdb_score,
  movie_cnt > params$movie_count)


```

###Countries by year, count of movies having rating more than _`r paste(params$imdb_score)`_ ###
```{r movielikes, message=FALSE, warning=FALSE}
plot_data %>% ggplot(aes(x = title_year,
                         y = imdb_score,
                         col = m_facebook_likes)) +
                         geom_point() +
                         geom_smooth(method = "loess",
                         alpha = 0.6,
                         size = 1)  +
                         labs(list(title = "IMDB Score by Facebook Likes", x = "Facebook Likes", y = "IMDB Score")) +
                         theme(plot.background = element_rect(fill = "#CFDBD3", color = "black", size =
                         3)) + theme(
                         panel.grid = element_blank(),
                         panel.background = element_blank(),
                         legend.key = element_blank(),
                         legend.background = element_blank(),
                         strip.background = element_blank()
                         ) +
                         
                         facet_grid(country ~ .)
```