---
title: "HR Efficiency Analysis"
date: "`r format(Sys.Date())`"
output: 
  html_document:
    keep_md: true
    fig_width: 14
    fig_height: 8
    out_width: '900px'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(caret)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(arm)
library(cluster)
library(GGally)
myTheme <- theme( panel.grid = element_blank(), legend.key = element_blank(), plot.title = element_text(hjust = 0.5))
```

## Problem
###Factors important to retain performing employees
* We will be analysing the data available in hand to identify avenues to improving hr efficiency.
* We will be building a predictive model to determine how long employee would stay and their probability in leaving.

## Data Source
https://www.kaggle.com/ludobenistant/hr-analytics

https://www.kaggle.com/ludobenistant/hr-analytics/downloads/human-resources-analytics.zip

#####Lets load the dataset
```{r Loading Data }
hr <- read.csv("./HR_comma_sep.csv", header = TRUE, stringsAsFactors = FALSE)
#hr_corrplot = hr
summary(hr)
```

* _satisfaction_level_ - Employee Level of satisfaction. It ranges from __0 (low satisfaction)__ to __1 (high satisfaction)__
* _last_evaluation_ - Latest performance evaluation rating of employee. It ranges from __0 (low rating)__ to __1 (high rating)__
* _number_project_ - Number of projects completed while at work. It varies from __2__ to __7__ projects, on an average an employee has worked on __4__ projects.
* _average_montly_hours_ - Average monthly work hours at workplace. It varies from __96 hrs__ to __310 hrs__ with a mean of __201 hrs__ per month.
* _time_spend_company_ - Number of years spent in the company. It ranges from __2__ years to __10__ years, with an average of around __4__ years until now or until they left.
* _Work_accident_ - Whether the employee had a workplace accident
* _left_ - Whether the employee left the workplace or not. __1__ means __left__ and __0__ means still with company. Probability of leaving is __0.23__.
* _promotion_last_5years_ - Whether the employee was promoted in the last five years. __1__ means promoted and __0__ means no promotion in last 5 years. Probability of promotion in last 5 years is __0.02__.


* _sales_ - Department in which they work for. _accounting_, _hr_, _IT_, _management_,  _marketing_, _product_mng_, _RandD_, _sales_, _support_, _technical_ are various departments employees belong to.
```{r summary_sales }
summary(as.factor(hr$sales))
```

* _salary_ - Salary as high, medium & low.
```{r summary_salary}
summary(as.factor(hr$salary))
```

* Fields _number_project_, _promotion_last_5years_, _left_, _Work_accident_, _sales_,_salary_ have discrete values.
* Looking at the summary of the data set there are no bad data , 'blanks', 'NA', null etc.

## Data Wrangling

Rename the _sales_ column to _dept_ and make columns _number_project_, _promotion_last_5years_, _left_, _Work_accident_, _sales_,_salary_ as factor
```{r}

hr$number_project <- as.factor(hr$number_project)
hr$promotion_last_5years <- as.factor(hr$promotion_last_5years)
hr$left <- as.factor(hr$left)
hr$Work_accident <- as.factor(hr$Work_accident)
hr$sales <- as.factor(hr$sales)
hr$salary <- as.factor(hr$salary)

names(hr)[9] <- "dept"

```

Lets look are structure
```{r}
  str(hr)
```

## Data Exploration

Lets look at correaltion between these various columns

```{r ggpairs_continous_attibutes}
hr %>%  dplyr::select(
  satisfaction_level,
  last_evaluation,
  average_montly_hours,
  number_project,
  left
  ) %>% ggpairs()
```

###__number_project__ 
* employees involved in 3-5 projects have high satisfaction level, less than 3 projects or more than 5 projects satisfaction is low
* last_evaluation increases with number_project
* average_monthly_hours also increases with number_project
* number of employees leaving (left=1) is high at 2 or less improves for 3 projects and then slowly creeps up until 5 projects and goes down
* Overall Looks like 3-5 projects would mean less employees quiting and good employee satisfaction.

###__average_monthly_hours__
* relation to satisfaction or last_evaluation is not very obvious
* average_monthly_hours seems to increase with number_project
* employees leaving (left=1)
  + 150 hours +/- 25 shows high number of employees leaving
	+ Again 225 hours and more employees leaving creaps to peak at around 250 hours and dips gradually to 300 hours
	+ 175 hours - 225 hours for average_monthly_hours seems to have low rate of employees leaving

###__last_evaluation__
* Between 0.4 to 0.6 evaluation rating, employees leaving (left=1) is high.
* It goes down at mid level and then picks up at around 0.75 and peaks around 0.85
* So low and higher end of evaluation rating employees leaving is high.
* Though one's leaving at lower rating is good for the company but the employees leaving having high rating is a concern for the company. 	
	
###__satisfaction_level__
* number projects between 3-5 show high satisfaction, less projects or too many projects tend to lower satisfaction
* employees leaving have lower satisfaction than employees not leaving the company

###__overall__
Looks like employees involved in 3 -5 projects and putting in 175 hours - 225 hours average monthly hours have lower rate of leaving company and have high satisfaction level.

```{r ggpairs_discrete_attributes}
hr %>%  dplyr::select(Work_accident, promotion_last_5years, salary, left) %>% ggpairs()

```

###__Work_accident__
* A small percent of employees involved in Work_accident are actually leaving company. Either Work_accident is not a very significant accident.
* Work_accident is higher at medium and low salary level than in higher salary level

###__promotion_last_5years__
* Very few employees are getting promoted
* number of employees getting promoted are even at different salary level.

###__salary__
* Higher number of employees leaving are in the medium & low salary level

###__overall__
Work_accident, promotion_last_5years & salary dont seem to have significant impact on employees leaving company.


Lets analyze _satisfaction_level_, _time_spend_company_, _last_evaluation_, _average_monthly_hours_ and _number_project_


```{r plot_data_exp_analysis}

satis_l <- hr %>% ggplot(aes(satisfaction_level)) +
  geom_histogram( binwidth = 0.05, aes(fill = left)) +
  labs(x = "satisfaction_level", y = "employees", title = "satisfaction level") + myTheme

nmbr_prj <- hr %>% ggplot(aes(as.numeric(number_project))) +
  geom_histogram( binwidth = 1, aes(fill = left)) +
  labs(x = "number_project", y = "employees", title = "Number of projects") + myTheme

lst_eval <- hr %>% ggplot(aes(last_evaluation)) +
  geom_histogram( binwidth = 0.05, aes(fill = left)) +
  labs(x = "last_evaluation", y = "employees", title = "Last evaluation") + myTheme

mnthly_hrs <- hr %>% ggplot(aes(average_montly_hours)) +
  geom_histogram( binwidth = 1, aes(fill = left)) +
  labs(x = "average_montly_hours", y = "employees", title = "Average montly hours") + myTheme

tm_spnd <- hr %>% ggplot(aes(time_spend_company)) +
  geom_histogram( binwidth = 1, aes(fill = left)) +
  labs(x = "time_spend_company", y = "employees", title = "Time Spend in Company") + myTheme


grid.arrange(satis_l, nmbr_prj, lst_eval, mnthly_hrs,tm_spnd,  nrow = 3)
```

* Using plots for __satisfaction_level__, __avaerage_monthly_hours__, __last_evaluation__ & __number_projects__:
  + Overall lower performing employees are leaving more. This warrants improvement in hiring process to avoid low performers
  + Aside from low performers, we can notice number of employees leaving creeeping up among mid to high performing. This is an area that needs to be also looked into for reduction in rate of attrition.  
  + Looks like employees involved in 3 -5 projects and putting in 175 hours - 225 hours average monthly hours have lower rate of leaving company and have high satisfaction level.


###Cluster Analysis 
```{r cluster1}
#lets identify optimum number of clusters
wssplot <- function(data, nc = 15, seed = 1234) {
  wss <- (nrow(data) - 1) * sum(apply(data, 2, var))
  for (i in 2:nc) {
    set.seed(seed)
    wss[i] <- sum(kmeans(data, centers = i)$withinss)
  }
  plot(1:nc,
       wss,
       type = "b",
       xlab = "Number of Clusters",
       ylab = "Within groups sum of squares")   
}

#lets consider employees who left the company
hr_clust <- hr %>% filter(left == 1)

#remove attribute left
hr_clust <- hr_clust[-7]

#lets make all fields numeric
hr_clust$dept <- as.numeric( as.factor(hr_clust$dept))
hr_clust$salary <- as.numeric( as.factor(hr_clust$salary))
hr_clust$number_project <- as.numeric( hr_clust$number_project)
hr_clust$Work_accident <- as.numeric( hr_clust$Work_accident)
hr_clust$promotion_last_5years <- as.numeric(hr_clust$promotion_last_5years)

#lets identify optimum number of clusters
wssplot(scale(hr_clust))
```

```{r clustering}
# lets consider 3 clusters 

fit.km <- kmeans(scale(hr_clust), 3)
clusplot(hr_clust,fit.km$cluster, color = TRUE)
```

```{r clustering_analysis_signifanct}
cluster_analysis <- cbind(hr_clust,fit.km$cluster)
names(cluster_analysis)[10] <- 'cluster'

satis_2 <- cluster_analysis %>% ggplot(aes(satisfaction_level)) + geom_histogram() + facet_grid(.~cluster)

num_prj2 <- cluster_analysis %>% ggplot(aes(number_project)) + geom_histogram() + facet_grid(.~cluster)

last_eval2 <- cluster_analysis %>% ggplot(aes(last_evaluation)) + geom_histogram() + facet_grid(.~cluster)

avg_hrs2 <- cluster_analysis %>% ggplot(aes(average_montly_hours)) + geom_histogram() + facet_grid(.~cluster)

grid.arrange(satis_2, num_prj2, last_eval2, avg_hrs2, nrow = 4)
```

###__cluster 1__
* Employees having overall higher satisfaction level, 
* involved 3-5 projects, 
* higher evaluation rating (0.8 -1) and 
* average monthly hours of 225  - 275 hours.

###__cluster 2__
* Employees having overall medium satisfaction level, 
* involved in less than 2 projects, 
* low to medium evaluation rating (<0.6) and 
* low average monthly hours of less than 175 hours.

###__cluster 3__
* Employees having overall low satisfaction level, 
* involved in more than 4 projects, 
* high evaluation rating (0.8-1) and 
* high average monthly hours of 225 - 350 hours.

###__overall__
* cluster 1 are employees that comapany need to find ways to retain.
* cluster 3 are over worked employees, company need to find ways to optimise work load and improve satisfaction level
* cluster 2 are under performing employees and company need to focus to minimize hiring employees of this category.

```{r cluster_analysis_not_imp}
promotion2 <- cluster_analysis %>% ggplot(aes(promotion_last_5years)) + geom_histogram() + facet_grid(.~cluster)
dept2 <- cluster_analysis %>% ggplot(aes(dept)) + geom_histogram() + facet_grid(.~cluster)
wrk_accdnt2 <- cluster_analysis %>% ggplot(aes(Work_accident)) + geom_histogram() + facet_grid(.~cluster)
sal2 <- cluster_analysis %>% ggplot(aes(salary)) + geom_histogram() + facet_grid(.~cluster)

grid.arrange(promotion2,dept2, wrk_accdnt2, sal2, nrow = 4)
```

###__overall__
All clusters have even distribution for promotion, department, work accident and salary. Not much significant observation that can be made.

##Regression Model

###Linear Regression
Lets build a model to determine how long an employee will stay using data from employees who already left


####Train Data
```{r}

set.seed(3456)
hr_left <- hr %>% filter(left==1)
trainIndex <- createDataPartition(hr_left$time_spend_company, p = .8, 
                                  list = FALSE, 
                                  times = 1)
head(trainIndex)

hrTrain <- hr_left[ trainIndex,]

```
###Test Data
```{r}
hrTest  <- hr_left[-trainIndex,]
```

###Models
####Lets use caret package to anlyze significant fields for linear model
```{r fields_lm}
plot(varImp(train(time_spend_company ~ ., data = hrTrain, method = "lm")))
```
Attributes satisfaction_level,number_project, avaerage_monthly_hours,promotion_last_5years, dept seems to be significant for linear regression
```{r}
lm_time_spend <- lm(time_spend_company~ satisfaction_level+number_project+ average_montly_hours+promotion_last_5years+ dept, data = hrTrain)
lm_time_spend_summary <- summary(lm_time_spend)
lm_time_spend_summary
```



_lm_time_spend_ has an R Squared of `r lm_time_spend_summary$r.squared` and adjusted R-Squared of `r lm_time_spend_summary$adj.r.squared`

#####R-Squared is good for the model.

###Lets predict using the model _lm_time_spend_
```{r}
predict_time_spend <- predict(lm_time_spend, newdata = hrTest)
#summary of prediction
summary(predict_time_spend)

#summary of actuals
summary(hrTest$time_spend_company)

```
The summary shows closeness in prediction to actuals.

###Lets measure the correlation

```{r compute_correlation_coef}
#correlation data
plot_data <- data.frame(predict_time_spend, hrTest$time_spend_company)
names(plot_data)[1] <- "predicted"
names(plot_data)[2] <- "actual"

#plot_data$predicted <- round(plot_data$predicted)

# plot predicted time spend vs actual time spend in conpany
plot_data %>% ggplot(aes(x = actual, y = predicted)) + geom_point() + geom_smooth(method = "lm")

correlation <- cor(plot_data)
correlation

#correlation
correlation[1,2]
```

The correlation coefficient is excellent __`r correlation[1,2]`__

```{r compute_rmse}
#lets see how good the model prediction was

#Sum of squared errors(SSE)
sse = sum((hrTest$time_spend_company - predict_time_spend ) ^ 2)
round(sse, digits = 2)

#Root mean squared errors (RMSE)
rmse = sqrt(sse / nrow(hrTest))
round(rmse, digits = 2)

```
####Linear model _lm_time_spend_ prediction has an 'Root mean squared errors' of `r round(rmse, digits = 2)` years and correlation coefficient of `r round(correlation[1,2], digits =2)`

###Logistic Regression
Lets build a model to predict if the employee will leave

####Train Data
```{r}

set.seed(3456)
trainIndex <- createDataPartition(hr$left, p = .8, 
                                  list = FALSE, 
                                  times = 1)
head(trainIndex)

hrTrain <- hr[ trainIndex,]

```
###Test Data
```{r}
hrTest  <- hr[-trainIndex,]
```

###Models

#####Significant fields

```{r log_reg_model_signi_fields}
plot(varImp(train(left ~ ., data = hrTrain, method = "bayesglm")))
```

#####Significant fields
* _satisfaction_level_, _Work_accident_, _salary_, _time_spend_company_, _number_project_, _average_montly_hours_, _promotion_last_5years_, _dept_

```{r log_model}
fitControl <- trainControl(method = "bayesglm",
                           number = 10,repeats = 10)
log_model <- train(left ~ satisfaction_level + Work_accident  + salary + time_spend_company + number_project + average_montly_hours + promotion_last_5years + dept, data = hrTrain, method = "bayesglm" )
glm_prediction_train <- predict(log_model, newdata = hrTrain)
glm_confustion_matrix_train <- confusionMatrix(glm_prediction_train, hrTrain$left, dnn = c("Predicted","actual"))
summary(log_model)
glm_confustion_matrix_train
```


```{r}
#Lets test prediction using test data 
testPredication <- predict(log_model,hrTest)
glm_confustion_matrix_test <- confusionMatrix(testPredication,hrTest$left, dnn = c("Predicted","actual"))

glm_confustion_matrix_test
```

```{r rf}
library(doMC)
registerDoMC(5)
hrTrain_1 <- hrTrain[createDataPartition(y = hrTrain$left,p = 0.3,list = FALSE),]
rf_model <- train(left~.,data = hrTrain_1,method = "rf",
                trControl = trainControl(method = "cv",number = 5),
                pallowParallel = TRUE)
print(rf_model)
rf_prediction_train = predict(rf_model, newdata = hrTrain)

rf_conf_matrix_train <- confusionMatrix(rf_prediction_train, hrTrain$left)
rf_conf_matrix_train

rf_prediction = predict(rf_model, newdata = hrTest)

rf_conf_matrix <- confusionMatrix(rf_prediction, hrTest$left)
rf_conf_matrix

```

#####Random Forest analysis has provided a very good model. 
* __Sensitivity__ : `r rf_conf_matrix$byClass["Sensitivity"]`
* __Specificity__ : `r rf_conf_matrix$byClass["Specificity"]`
* __Accuracy__    : `r rf_conf_matrix$overall["Accuracy"]`
* __Kappa__       : `r rf_conf_matrix$overall["Kappa"]`

```{r reg_summary}
overall <- cbind(
  glm_confustion_matrix_train$overall,
  glm_confustion_matrix_test$overall,
  rf_conf_matrix_train$overall,
  rf_conf_matrix$overall
  )
  
byClass <- cbind(
  glm_confustion_matrix_train$byClass,
  glm_confustion_matrix_test$byClass,
  rf_conf_matrix_train$byClass,
  rf_conf_matrix$byClass
  )
  
all_conf_matrix <- rbind(overall, byClass)
all_conf_matrix <- data.frame(all_conf_matrix)
names(all_conf_matrix) <- c("glm_train", "glm_test", "rf_train", "rf_test")

all_conf_matrix$glm_train <- round(all_conf_matrix$glm_train, digits = 4)
all_conf_matrix$glm_test <- round(all_conf_matrix$glm_test, digits = 4)
all_conf_matrix$rf_train <- round(all_conf_matrix$rf_train, digits = 4)
all_conf_matrix$rf_test <- round(all_conf_matrix$rf_test, digits = 4)

all_conf_matrix

```




